---
title: "intro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 7,
  warning = FALSE,
  errors = FALSE,
  messages = FALSE
)
```

## Description

`bamExtras` is an R package that contains functions and data to support
stock assessments of fish stocks in the Southeast US Atlantic using the Beaufort
Assessment Model (BAM). It provides support and organization for data and functions used to run stock assessments and diagnostics with BAM. Currently it contains all of the most recent versions of BAM for 14 stocks and can be used to modify and run BAM, if ADMB is installed. It also contains various support and plotting functions


## Example usage

### Load the package

First, install and load the package. 

```{r, echo=TRUE, message=FALSE}
# Install and load package
#devtools::install_github("nikolaifish/bamExtras")
library(bamExtras)
# Load dependencies
library(stringr)
```

### Data
`bamExtras` contains BAM input files: `dat` (data), `tpl` (template model code), `cxx` (convert C++ output to R objects), and `rdat` (R output) files for the most recent models. The `dat`, `tpl`, and `cxx` files are stores as character vectors with each item representing a line of the original file. Other files like "admb2r.cpp" and "cleanup.bat" are also stored in `bamExtras` for running BAM.

```{r, echo=TRUE}
# see all data objects in bamExtras
data(package="bamExtras")$result[,"Item"]
```

Since different naming conventions are sometimes used in BAM, the function `standardize_rdat()` can be used to modify names to follow common naming conventions. Common naming conventions are important for some functions in `bamExtras` and allow for general analysis.

```{r, echo=TRUE, message=FALSE}
rdat_BS <- standardize_rdat(rdat_BlackSeaBass)
rdat_RP <- standardize_rdat(rdat_RedPorgy)
```

We can simply use the stored data to plot results from an assessment

```{r}
par(mfrow=c(3,2),mar=c(2,2,1,1),mgp=c(1,0.2,0),tck=-0.01)
with(rdat_BS$t.series,{
     plot(year,N,type="o")
     plot(year,recruits,type="o")  
     plot(year,SSB,type="o")
     plot(year,F.full,type="o")
     plot(year,SSB.msst,type="o")
     plot(year,F.Fmsy,type="o")
  })
```

We can plot results from multiple assessments

```{r}
spcol <- c("BS"="black","RP"="red")

par(mfrow=c(3,2),mar=c(2,2,1,1),mgp=c(1,0.2,0),tck=-0.01)
ylim <- range(c(0,rdat_BS$t.series$N,rdat_RP$t.series$N),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,N,type="o",col=spcol["BS"],ylim=ylim))
with(rdat_RP$t.series,points(year,N,type="o",col=spcol["RP"]))

legend("top",legend=c("Black Sea Bass","Red Porgy"),
       pch=1,col=spcol,lty=1)

ylim <- range(c(0,rdat_BS$t.series$recruits,rdat_RP$t.series$recruits),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,recruits,type="o",col=spcol["BS"],ylim=ylim))
with(rdat_RP$t.series,points(year,recruits,type="o",col=spcol["RP"]))

ylim <- range(c(0,rdat_BS$t.series$SSB,rdat_RP$t.series$SSB),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,SSB,type="o",col=spcol["BS"],ylim=ylim))
with(rdat_RP$t.series,points(year,SSB,type="o",col=spcol["RP"]))

ylim <- range(c(0,rdat_BS$t.series$F.full,rdat_RP$t.series$F.full),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,F.full,type="o",col=spcol["BS"],ylim=ylim))
with(rdat_RP$t.series,points(year,F.full,type="o",col=spcol["RP"]))

ylim <- range(c(0,rdat_BS$t.series$SSB.msst,rdat_RP$t.series$SSB.msst),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,SSB.msst,type="o",col=spcol["BS"],ylim=ylim))
with(rdat_RP$t.series,points(year,SSB.msst,type="o",col=spcol["RP"]))

ylim <- range(c(0,rdat_BS$t.series$F.Fmsy,rdat_RP$t.series$F.Fmsy),na.rm=TRUE)
with(rdat_BS$t.series,plot(year,F.Fmsy,type="o",col="black",ylim=ylim))
with(rdat_RP$t.series,points(year,F.Fmsy,type="o",col=spcol["RP"]))
```

## Functions

One of core functions in this package is `bam2R`. is a helpful function compares corresponding `dat` and `tpl` files from an assessment, and builds an R list `L_init` of all the `init` objects initialized in the `DATA_SECTION` of the `tpl`. It then combines `L_init` and the BAM files `dat`, `tpl`, and `cxx` into an output list.

```{r, echo=TRUE}
bam_RP <- bam2r(dat_obj=dat_RedPorgy,tpl_obj=tpl_RedPorgy,cxx_obj=cxx_RedPorgy)
```

Drawing upon data objects stored in the package, `bam2R` can also be used in a simpler way with the `CommonName` argument

```{r}
bam_RP <- bam2r("RedPorgy")
```

## Running BAM
If ADMB is installed, we can then rerun the stock assessment. The code in this section is provided by commented out for the vignette.

```{r}
# Run a bam model and assign rdat output to object
#rdat_RP <- run_bam(bam=bam_RP,fileName="RP")
```

And `bam2r` lets us modify the data inputs. Then we can rerun BAM.


```{r}
# Modify data input from a BAM model and incorporate it back into the dat file object
# L_init2 <- bam_RP$L_init
# L_init2$set_steep[c(1,5)] <- paste(0.6) # Change steepness
# L_init2$set_steep[4] <- paste(-abs(as.numeric(L_init2$set_steep[4]))) # Fix steepness
# bam_RP2 <- bam2r("RedPorgy",L_init_user=L_init2)
# rdat_RP2 <- run_bam(bam=bam_RP2,fileName="RP2")
```

Then we can compare the two models

```{r}
# spcol <- c("RP"="black","RP2"="red")
# 
# par(mfrow=c(3,2),mar=c(2,2,1,1),mgp=c(1,0.2,0),tck=-0.01)
# ylim <- range(c(0,rdat_RP$t.series$N,rdat_RP2$t.series$N),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,N,type="o",col=spcol["RP"],ylim=ylim))
# with(rdat_RP2$t.series,points(year,N,type="o",col=spcol["RP2"]))
# 
# legend("top",legend=c("Red Porgy","Red Porgy2"),
#        pch=1,col=spcol,lty=1)
# 
# ylim <- range(c(0,rdat_RP$t.series$recruits,rdat_RP2$t.series$recruits),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,recruits,type="o",col=spcol["RP"],ylim=ylim))
# with(rdat_RP2$t.series,points(year,recruits,type="o",col=spcol["RP2"]))
# 
# ylim <- range(c(0,rdat_RP$t.series$SSB,rdat_RP2$t.series$SSB),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,SSB,type="o",col=spcol["RP"],ylim=ylim))
# with(rdat_RP2$t.series,points(year,SSB,type="o",col=spcol["RP2"]))
# 
# ylim <- range(c(0,rdat_RP$t.series$F.full,rdat_RP2$t.series$F.full),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,F.full,type="o",col=spcol["RP"],ylim=ylim))
# with(rdat_RP2$t.series,points(year,F.full,type="o",col=spcol["RP2"]))
# 
# ylim <- range(c(0,rdat_RP$t.series$SSB.msst,rdat_RP2$t.series$SSB.msst),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,SSB.msst,type="o",col=spcol["RP"],ylim=ylim))
# with(rdat_RP2$t.series,points(year,SSB.msst,type="o",col=spcol["RP2"]))
# 
# ylim <- range(c(0,rdat_RP$t.series$F.Fmsy,rdat_RP2$t.series$F.Fmsy),na.rm=TRUE)
# with(rdat_RP$t.series,plot(year,F.Fmsy,type="o",col="black",ylim=ylim))
# with(rdat_RP2$t.series,points(year,F.Fmsy,type="o",col=spcol["RP2"]))
```


## Conduct a catch curve analysis

```{r}
# read in black sea bass age compositions used in the recent stock assessment
rdat <- rdat_BlackSeaBass
comp <- rdat$comp.mats # List of composition matrices
acomp <- comp[grepl("^acomp.*.ob$",names(comp))] # list of observed age composition matrices
nm_comp_i <- names(which.max(lapply(acomp,nrow))) # Get the name of the age comps with the most years of data
comp_i <- acomp[[nm_comp_i]]
cc_i <- catch_curve(comp_i)
# Plot comps and add catch curves (note the log-tranformation of the composition data)
comp_plot(log(comp_i),cc=cc_i,fillComp = FALSE,ylab= "log(proportion)",xlab="age",title=paste("catch curves for",nm_comp_i))
# Plot Z estimates over time for multiple sets of age compositions
par(mfrow=c(1,1),mgp=c(1,.2,0),mar=c(2,2,1,1),oma=c(0,0,0,0),tck=0.01)
M <- rdat$parms$M.constant
cc <- lapply(acomp,catch_curve)
abb <- gsub("^(acomp.)|(.ob)$","",names(cc)) # abbreviations of age composition fleets
cols <- rainbow(length(cc))
xlim <- range(unlist(lapply(cc,function(x){as.numeric(x$year)})),na.rm=TRUE)
ylim <- range(c(0,unlist(lapply(cc,function(x){-as.numeric(x$slope)}))),na.rm=TRUE)
for(i in 1:length(cc)){
  xi <- cc[[i]]
  year <- as.numeric(xi$year)
  Z <- -xi$slope
  if(i==1){
    plot(year,Z,xlim=xlim,ylim=ylim,type="o",col=cols[i])
  }else{
    points(year,Z,type="o",col=cols[i])
  }
}
# Add reference line for natural mortality
abline(h=M,lty=2,lwd=2)
text(par("usr")[1]+0.05*diff(par("usr")[1:2]),M,labels="M",pos=3)
legend("bottomleft",legend=abb,col=cols,pch=1,lty=1,bty="n")
```
