---
title: "Calculating reference points"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculating reference points}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height= 7
)
```

```{r setup, echo=FALSE,message=FALSE}
library(bamExtras)
```

# Introduction

The purpose of this vignette is to demonstrate how reference points can be calculated using thr `bamExtras::F_calc` function. There are a couple of examples in the `F_calc` help file, but this vignette provides examples for all stock assessments loaded into the `bamExtras` package.

## Atlantic Menhaden

```{r}
rdat <- rdat_AtlanticMenhaden
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wholewgt.wgted.L.mt,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=3,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40
```

## Black Sea Bass

```{r}
rdat <- rdat_BlackSeaBass
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=6,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$Mean.biascorr*pr$Mean.R0,
       Px=0.5
       )
F50 <- c(bam=rdat$spr.brps$F50,bamExtras=Fcout$RefPts$F_Px)
F50
```

## Cobia

```{r}
rdat <- rdat_Cobia
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$steep,
       R0=pr$R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=3,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$Mean.biascorr*pr$Mean.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40
```

## Gag

```{r}
rdat <- rdat_GagGrouper
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb.gut,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy
```

## Gray triggerfish

```{r}
rdat <- rdat_GrayTriggerfish
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$Mean.biascorr*pr$Mean.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40
```

## Greater amberjack

```{r}
rdat <- rdat_GreaterAmberjack
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

## Red grouper

```{r}
rdat <- rdat_RedGrouper
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

## Red porgy

```{r}
rdat <- rdat_RedPorgy
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

## Red Snapper

```{r }
# When there is no SRR, the function will still compute SPR without computing MSY-based equilibrium values
rdat <- rdat_RedSnapper
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(age=as$age,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wholewgt.wgted.L.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=rdat$parms$spawn.time,
       plot_digits=4,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$Mean.biascorr*pr$Mean.R0,
       F_n=1001
       )

F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40
# 
# Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
# Fmsy
```

## Scamp Grouper

```{r}
rdat <- rdat_ScampGrouper
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=0.6,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$Mean.biascorr*pr$Mean.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40
# 
# Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
# Fmsy

```

## Snowy Grouper

```{r}
rdat <- rdat_SnowyGrouper
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=0.6,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

## Spanish Mackerel

```{r}
rdat <- rdat_SpanishMackerel
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wholewgt.wgted.L.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

## Tilefish

```{r}
rdat <- rdat_Tilefish
pr <- rdat$parms
as <- rdat$a.series

Fcout <- F_calc(
       age=as$age,
       rec_sigma=pr$R.sigma.par,
       h=pr$BH.steep,
       R0=pr$BH.R0,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wgt.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=pr$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001,
       N_ast_ob=as.numeric(tail(rdat$N.age.spawn,1)),
       R_neq=pr$BH.biascorr*pr$BH.R0,
       Px=0.4
       )
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```


## Vermilion Snapper

```{r}
rdat <- rdat_VermilionSnapper
pr <- rdat$parms
as <- rdat$a.series

# pdout <- pop_demo(
#        Fi=pr$F30,
#        age=as$age,
#        h=pr$BH.steep,
#        R0=pr$R0,
#        rec_sigma=rdat$parms$R.sigma.par,
#        M=as$M,
#        sel_L=rdat$sel.age$sel.v.wgted.L,
#        sel_D=rdat$sel.age$sel.v.wgted.D,
#        PS=rep(1,length(as$age)),
#        w=as$wholewgt.wgted.L.klb,
#        ep=as$reprod,
#        plots=TRUE,
#        P_st=rdat$parms$spawn.time,
#        plot_digits=4)
# pdout

Fcout <- F_calc(age=as$age,
       h=pr$BH.steep,
       R0=pr$R0,
       rec_sigma=rdat$parms$R.sigma.par,
       M=as$M,
       sel_L=rdat$sel.age$sel.v.wgted.L,
       sel_D=rdat$sel.age$sel.v.wgted.D,
       PS=rep(1,length(as$age)),
       w=as$wholewgt.wgted.L.klb,
       ep=as$reprod,
       plots=TRUE,
       P_st=rdat$parms$spawn.time,
       plot_digits=4,
       Fmax=1,
       F_n=1001)
F40 <- c(bam=rdat$spr.brps$F40,bamExtras=Fcout$RefPts$F_Px)
F40

Fmsy <- c(bam=pr$Fmsy,bamExtras=Fcout$RefPts$Fmsy)
Fmsy

```

